---
output: html_document
---
## Data Cleaning
To clean the data, the first row index and all the NA columns were removed colomuns with NA were removed. The training and testing data were saved as pml-training2.csv and pml-testing2.csv 
```{r}
# Clear current directory
rm(list = ls())


# Read cleaned training and testing data 
training <- read.table(file = "pml-training2.csv", 
                       header = TRUE, sep = ",", quote = "")
testing <- read.table(file = "pml-testing2.csv", 
                      header = TRUE, sep = ",", quote = "")


# Change the numeric type to integer type to ensure same data type in training as well as testing data set.
training$magnet_dumbbell_z <- as.integer(training$magnet_dumbbell_z)
training$magnet_forearm_y <- as.integer(training$magnet_forearm_y)
training$magnet_forearm_z <- as.integer(training$magnet_forearm_z)
# Change the 
levels(testing$new_window) <- levels(training$new_window)
```


## Exploratory Analysis of Data

Cross Validation was performed to find out sample errors. 

```{r}
# Install randomForest package
#install.packages("randomForest")
library(randomForest)
#install.packages("caret")
#install.packages('e1071', dependencies=TRUE)
library(caret)
```

## Exploratory Analysis of Data
```{r}
set.seed(789)
# Define cross-validation
fitControl = trainControl( method = "cv", number = 2)
# Perform cross validation
crossvalid <- train(classe ~ ., data = training, method = "rf", 
  trControl = fitControl)
crossvalid$bestTune$mtry
```

## Exploratory Analysis of Data
```{r}
crossvalid
```

## Build random forest model using Training Set
Number of variables randomly sampled is: `r crossvalid$bestTune$mtry`
```{r}
RandomForest = randomForest(classe ~ ., data = training,mtry = crossvalid$bestTune$mtry)
PredictForTrain = predict(RandomForest)
table(PredictForTrain, training$classe)
```

## Predict Test Data
```{r}
PredictForest = predict(RandomForest, newdata = testing)
PredictForest
```
  Print Predictions to a file.
```{r}
# Function to write a vector to files
pml_write_files = function(x){
  n = length(x)
  for(i in 1:n){
    filename = paste0("problem_id_", i ,".txt")
    write.table(x[i], file = filename, quote = FALSE,
                row.names = FALSE, col.names = FALSE)
  }
}
pml_write_files(PredictForest)

```


